\thetitle

### 2021-03-02 writing/introduction.tex
touching up the intro

Theoreticians of gradual typing have focused on blame theorems from the very
beginning~\cite{mf-toplas-2009, tf-dls-2006}. ``Well-typed
[components]\footnote{The original authors got this word wrong.} can't be
blamed'' turned the theorem into a slogan~\cite{wf-esop-2009}. Academic systems
(Reticulated Python~\cite{vsc-dls-2019, vss-popl-2017, vksb-dls-2014} and Typed
Racket~\cite{tf-dls-2006,tf-popl-2008,tfffgksst-snapl-2017,tf-icfp-2010}) come
with sophisticated blame assignment strategies. Their academic creators embrace
the idea that blame can help practicing programmers find impedance mismatches
between the types they added to a software system and the behavior of the
remaining untyped components.

@@

This paper's {\em first contribution\/} is {\em a method for evaluating the
effectiveness of blame assignment strategies\/} in the gradual typing world.
The top-level innovation is the idea of a {\em rational programmer\/}, that is,
a programmer that acts only in response to available information (see
sec.~\ref{sec:why-rational}). In the case of an impedance mismatch, the
available information consists of the error message and the current state of the
program. The rational programmer can hence use the former to change the
latter---and this systematic, information-driven process can be
implemented and tested at scale on real programs.\footnote{\citet{cc-snapl-19} use this tracing method to localize
type errors in inference-based gradual typing systems.}  Testing it systematically requires
overcoming major challenges: injecting representative impedance mismatches;
putting the various kinds of error information to comparable use; and sampling
the huge space of possibilities (see sec.~\ref{sec:challenges}).

The paper's {\it second contribution\/} is a set of {\em results from applying
the evaluation method to three distinct blame assignment strategies in
approximately 72,200 different scenarios\/} (see sec.~\ref{sec:results}): (1)~{\it
Transient\/}, i.e. Reticulated's tracking of typed/untyped boundaries~\cite{vss-popl-2017}; (2)~{\it
Natural\/}, i.e. the use of a higher-order contract system and its blame
assignment~\cite{ff-icfp-2002}; and (3) {\it Erasure\/}, i.e. the approach of
industrial systems, which forgo blame in favor of error messages from the safety
checks of the underlying language. The results (see
sec.~\ref{sec:discussion}) are at least somewhat surprising.  In principle
they validate the conjectures behind the work of theoreticians.  First, a good
blame assignment strategy helps with the search for impedance mismatches between
the specified types and the behavior of untyped components.  Second, Natural's
wrapper-based blame tracking is more useful than Transient's ``collective
blame'' tracking algorithm, which in turn is superior to Erasure. {\em But\/},
the application of the method also indicates problems with the expectations of
theoreticians. The first problem is that the cost of Transient's blame can be
huge. The second one concerns the difference between the methods; neither is
{\it Natural\/} vastly better than {\it Transient\/} nor are the two blame
assignment methods clearly preferable to {\it Erasure\/}.  In turn, these
problems suggests that, on one hand, the existing theory does not predict
practice properly, and on the other hand, the existing practice may need to find
additional ways to collect data.

>>>>

Theoreticians of gradual typing have focused on blame theorems from the very
beginning~\cite{mf-toplas-2009, tf-dls-2006}. ``Well-typed
[components]\footnote{The original authors got this word wrong.} can't be
blamed'' turned the theorem into a slogan~\cite{wf-esop-2009}. Academic systems
(Reticulated Python~\cite{vsc-dls-2019, vss-popl-2017, vksb-dls-2014} and Typed
Racket~\cite{tf-dls-2006,tf-popl-2008,tfffgksst-snapl-2017,tf-icfp-2010}) come
with sophisticated checking and blame assignment strategies (see sec.~\ref{sec:landscape}). Their academic creators embrace
the idea that blame can help practicing programmers find impedance mismatches
between the types they added to a software system and the behavior of the
remaining untyped components.

@@

This paper's {\em first contribution\/} is {\em a method for evaluating the
effectiveness of blame assignment strategies\/} in the gradual typing world.
The top-level innovation is the idea of a {\em rational programmer\/}, that is,
a programmer that acts only in response to available information (see
sec.~\ref{sec:why-rational}). In the case of an impedance mismatch, the
available information consists of the error message and the current state of the
program. The rational programmer can hence use the former to change the
latter---and this systematic, information-driven process can be implemented and
tested at scale on real programs.  Turning this idea into a scientific
experiment requires overcoming major challenges: injecting representative
impedance mismatches; putting the various kinds of error information to
comparable use; and sampling the huge space of possibilities (see
sec.~\ref{sec:challenges} and, for details, secs.~\ref{sec:rational} through~\ref{sec:sample}).

The paper's {\it second contribution\/} is a set of {\em results from applying
the evaluation method to three distinct blame assignment strategies in
approximately 72,200 different scenarios\/} (see sec.~\ref{sec:results}): (1)~{\it
Transient\/}, i.e. Reticulated's tracking of typed/untyped boundaries~\cite{vss-popl-2017}; (2)~{\it
Natural\/}, i.e. the use of a higher-order contract system and its blame
assignment~\cite{ff-icfp-2002}; and (3) {\it Erasure\/}, i.e. the approach of
industrial systems, which forgo blame in favor of error messages from the safety
checks of the underlying language. The results (see
sec.~\ref{sec:discussion}) are at least somewhat surprising.  In principle
they validate the conjectures behind the work of theoreticians.  First, a good
blame assignment strategy helps with the search for impedance mismatches between
the specified types and the behavior of untyped components.  Second, Natural's
wrapper-based blame tracking is more useful than Transient's ``collective
blame'' tracking algorithm, which in turn is superior to Erasure. {\em But\/},
the application of the method also indicates problems with the expectations of
theoreticians. The first problem is that the cost of Transient's blame can be
huge. The second one concerns the difference between the methods; neither is
{\it Natural\/} vastly better than {\it Transient\/} nor are the two blame
assignment methods clearly preferable to {\it Erasure\/}.  In turn, these
problems suggests that, on one hand, the existing theory does not predict
practice properly, and on the other hand, the existing practice may need to
think of additional experiments to collect more data.
